---
title: "Geophy_NYC_Prelimnary_Analysis"
author: "Deepak Sharma"
date: "March 15, 2018"
output: html_document
---

Load libraries
```{r,include=FALSE}
x <- c("data.table","ggmap","dplyr","tidyr","rgdal")
lapply(x, library, character.only = TRUE)

```


Download the Condominium and Garden dataset from the website and check their dimensions

```{r,echo=FALSE}
Condominium <- fread("https://data.cityofnewyork.us/api/views/ikqj-pyhc/rows.csv?accessType=DOWNLOAD")
Gardens <- fread("https://data.cityofnewyork.us/api/views/ajxm-kzmj/rows.csv?accessType=DOWNLOAD")
dim(Condominium) #1165x46
dim(Gardens) #536x17
```


Look at the two data files structures

```{r,results="hide"}
#Condominium dataset
glimpse(Condominium)

#Gardens dataset
glimpse(Gardens)
```

Remove NA's from the columns Latitude,Longitude,Postcode from both data sets and rename the datsets

```{r}
condominium <- Condominium %>% drop_na(Latitude,Longitude,Postcode) #New dimension 1148x46
gardens <-   Gardens %>% drop_na(Latitude,Longitude,Postcode) #New dimension 429x17
```

Explore some on the columns in condominium data to get idea about range of values

```{r}
names(condominium)
hist(condominium$`MANHATTAN CONDOMINIUM PROPERTY Gross Income per SqFt`,col="green")
hist(condominium$`MANHATTAN CONDOMINIUM PROPERTY Year Built`,col="green")
hist(condominium$`MANHATTAN CONDOMINIUM PROPERTY Market Value per SqFt`,col="green")
hist(condominium$`MANHATTAN CONDOMINIUM PROPERTY Gross SqFt`,col="green") #Two properties are quite big.
hist(condominium$`MANHATTAN CONDOMINIUM PROPERTY Total Units`,col="green")

```

Every thing looks normal till now
#Observations:
#Two of the properties are much larger than the others.Their addresses are:
#Adress1: 333 East 30 Street - No Park - Gross income/sqft = 28
#Adress2: 1601 3 avenue - Has Ruppert park - Gross income/sqft = 37

Look at the two large properties:

```{r,include=FALSE}
Very_large_GrossSqFt <- filter(condominium,`MANHATTAN CONDOMINIUM PROPERTY Gross SqFt`>1000000)
glimpse(Very_large_GrossSqFt)
```

TASK 1 - (a) Visualize garden datset on the map

```{r}
mapC <- get_map("New York City",zoom=11)
#Garden distribution across the city
ggmap(mapC)+geom_point(data=gardens,aes(x=Longitude,y=Latitude))
```

TASK 1 - (b) Visualize condominium datset on the map

```{r}
mapM <- get_map("Manhattan,New York City",zoom=12)
ggmap(mapM)+geom_point(data=condominium,aes(x=Longitude,y=Latitude))

ggmap(mapM)+geom_point(data=condominium,aes(x=Longitude,y=Latitude,
color=`MANHATTAN CONDOMINIUM PROPERTY Market Value per SqFt`),na.rm = T)+
scale_color_gradient(low="beige",high="blue")

ggmap(mapM)+geom_point(data=condominium,aes(x=Longitude,y=Latitude,
color=`MANHATTAN CONDOMINIUM PROPERTY Gross Income per SqFt`),na.rm = T)+
scale_color_gradient(low="beige",high="blue")

ggmap(mapM)+geom_point(data=condominium,aes(x=Longitude,y=Latitude,
color=`MANHATTAN CONDOMINIUM PROPERTY Gross SqFt`),na.rm = T)+
scale_color_gradient(low="beige",high="blue")

ggmap(mapM)+geom_point(data=condominium,aes(x=Longitude,y=Latitude,
color=`MANHATTAN CONDOMINIUM PROPERTY Year Built`),na.rm = T)+
scale_color_gradient(low="beige",high="blue")
#******************************************************************************
```

TASK 2 
 
First we generate a new table NTA_info with only few of the important features,
and arrange it using "MANHATTAN CONDOMINIUM PROPERTY Gross Income per SqFt"

#NTA_info table
```{r}
NTA_info <- condominium %>% group_by(NTA) %>% 
  summarize(Mean_property_value_perSqFt = mean(`MANHATTAN CONDOMINIUM PROPERTY Market Value per SqFt`,na.rm=TRUE),
  Mean_income_perSqFt = mean(`MANHATTAN CONDOMINIUM PROPERTY Gross Income per SqFt`,na.rm=TRUE),
  Mean_area_perSqFt = mean(`MANHATTAN CONDOMINIUM PROPERTY Gross SqFt`,na.rm=TRUE),
  Mean_year_built = mean(`MANHATTAN CONDOMINIUM PROPERTY Year Built`,na.rm=TRUE)) %>% 
  arrange(desc(Mean_income_perSqFt))

glimpse(NTA_info)
```

TASK - 3 

Creating Postcode_garden table based on total number of gardens in each Post code

```{r}
Postcode_garden <- gardens %>% group_by(Postcode) %>% summarise(Number_of_gardens=n()) %>% arrange(desc(Number_of_gardens))
glimpse(Postcode_garden) #dimension 81x2
```

Creating Postcode_condo table with several parameters grouped by Postcode values

```{r}
Postcode_condo <- condominium %>% group_by(Postcode) %>% 
  summarize(Mean_property_value_perSqFt = mean(`MANHATTAN CONDOMINIUM PROPERTY Market Value per SqFt`,na.rm=TRUE),
  Mean_income_perSqFt = mean(`MANHATTAN CONDOMINIUM PROPERTY Gross Income per SqFt`,na.rm=TRUE),
  Mean_area_perSqFt = mean(`MANHATTAN CONDOMINIUM PROPERTY Gross SqFt`,na.rm=TRUE),
  Mean_year_built = mean(`MANHATTAN CONDOMINIUM PROPERTY Year Built`,na.rm=TRUE)) %>% 
  arrange(desc(Mean_property_value_perSqFt))
glimpse(Postcode_condo) #dimension 44x5
```


Merging Postcode_garden and Postcode_condo in one table - Postcode_info
#Postcode_info table 

```{r}
#Combining Postcode_garden and Postcode_condo tables.
Postcode_info <- full_join(Postcode_garden,Postcode_condo) %>% arrange(Postcode) #101x6
glimpse(Postcode_info)
```

Scatter plot and correlation matrix
```{r}
plot(Postcode_info)
Postcode_matrix =as.matrix(Postcode_info)
cor(Postcode_matrix, use = "pairwise.complete.obs")
```

#Read attached PDF report.
